---
title: "Timeseries analysis"
author: "Katherine Mueller"
date: "2/12/2021"
output: html_document
---
```{r}
library(janitor)
library(dplyr)
library(tidyr)
library(phyloseq)
```

Set file locations for count and key files
```{r}
subjectA_counts_file <- "C:/Users/katmu/Dropbox/Valdivia\ Lab/Projects/LAD\ Timeseries/Original\ Data/share/microbiota/subjectA.gut.M"
subjectA_key_file <- "C:/Users/katmu/Dropbox/Valdivia\ Lab/Projects/LAD\ Timeseries/Original\ Data/share/microbiota/subjectA.gut.key"
subjectB_counts_file <- "C:/Users/katmu/Dropbox/Valdivia\ Lab/Projects/LAD\ Timeseries/Original\ Data/share/microbiota/subjectB.gut.M"
subjectB_key_file <- "C:/Users/katmu/Dropbox/Valdivia\ Lab/Projects/LAD\ Timeseries/Original\ Data/share/microbiota/subjectB.gut.key"
```

Open count and key files as dataframes
```{r}
subjectA_counts <- read.delim(subjectA_counts_file, sep = " ", header = FALSE)
subjectA_key <- read.delim(subjectA_key_file, sep = "\t", header = FALSE)
subjectB_counts <- read.delim(subjectB_counts_file, sep = " ", header = FALSE)
subjectB_key <- read.delim(subjectB_key_file, sep = "\t", header = FALSE)
```

Turn count dataframes into OTU format for phyloseq. Because the taxa assignments are the same, I will combine subject A and subject B into one dataframe.
```{r}
subjectA_counts_t <- t(subjectA_counts)
subjectA_counts_t <- row_to_names(subjectA_counts_t, 1)
colnames(subjectA_counts_t) <- paste("A", colnames(subjectA_counts_t), sep = "_")
rownames(subjectA_counts_t) <- subjectA_key$V1

subjectB_counts_t <- t(subjectB_counts)
subjectB_counts_t <- row_to_names(subjectB_counts_t, 1)
colnames(subjectB_counts_t) <- paste("B", colnames(subjectB_counts_t), sep = "_")
rownames(subjectB_counts_t) <- subjectB_key$V1

counts_table <- merge(subjectA_counts_t, subjectB_counts_t, by = 0, all = TRUE)
#remove timepoints with no sequencing results
emptycols <- sapply(counts_table, function (x) all(is.na(x)))
counts_table <- counts_table[!emptycols]

otus <- otu_table(counts_table, taxa_are_rows = TRUE)
```

Turn key files into taxa format for phyloseq. The OTU identifiers and taxa assignments are the same between both key files, so I only need one.
```{r}
subjectA_taxa <- subjectA_key
row.names(subjectA_taxa) <- subjectA_taxa$V1
subjectA_taxa$V3 <- substring(subjectA_taxa$V3, 4)
subjectA_taxa <- subjectA_taxa %>%
  select(V3) %>%
  separate(V3, c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";.__")
subjectA_tax_table <- subjectA_taxa %>%
  as.matrix(subjectA_taxa) %>%
  tax_table()
```

Make phyloseq objects
```{r}
subjectA_ps <- phyloseq(subjectA_otus, subjectA_tax_table)
subjectA_ps
```
